Towards Reproducibility in ANS Performance

### 2019-10-04 paper.Rmd
final code version

---
title: "Busting the Myth! - Reproducibility in ANS Performance"
author:
- Enrico Spinielli
- Rainer Koelle
date: "`r format(Sys.time(), '%a %b %d %Y @  %X')`"
output:
  bookdown::pdf_book:
    base_format: rticles::ieee_article
    fig_caption: no
    includes:
      in_header: header.tex
      keep_tex: yes
    number_sections: yes
    toc: no
    with_dblfloatfix: no
    with_subfig: no
  html_document: default
bibliography: references.bib
color-links: yes
csl: ieee-transactions-on-aerospace-and-electronic-systems.csl
affiliation:
  institution-columnar: true
  institution:
  - author:
    - email: enrico.spinielli@eurocontrol.int
      name: Enrico Spinielli
    - email: rainer.koelle@eurocontrol.int
      name: Rainer Koelle
    department: Performance Review Unit
    location: Brussels, Belgium
    name: EUROCONTROL
keywords:
- reproducibility
- ANS performance
- flight trajectory
- open data
- performance indicator
link-citations: yes
author-meta: Enrico Spinielli, Rainer Koelle
papersize: a4paper
abstract: "Air transportation is undergoing a fundamental transformation. On the political,
  strategic, and operational level work is directed to meet the future growth of air
  traffic.\nTo meet this goal, operational concepts and technological enablers are
  deployed with the promises of distinct performance improvements. These improvements
  are typically widely marketed, however, the underlying evidence is not made available.
  From that respect Air Navigation System Performance faces varying levels of transparency
  as seen in other industries. \nThis paper implements the reproducibility paradigm
  by providing data, code, and results openly to enable another researcher or analyst
  to reproduce - and potentially - improve or build on our work. This work is part
  of a wider thread to establish Open Data and Open Source based data analytical capabilities
  for Air Navigation System Performance in Europe.\nThe implementation of the paradigm
  is demonstrated on the basis of the performance reference trajectory by replicating
  a classical performance measure, and then moving into investigating the application
  of trajectory based information for enhancing the current state of performance monitoring
  for the arrival phase. This paper applies the concept and approach as an use case
  analysis for 2 European airports with different strategies for sequencing arrival
  traffic (i.e. stack holding and point merge).\nThe work reported demonstrates the
  feasibility of the approach and reflects a stepping stone in the roll-out of the
  performance reference trajectory within the European context. The work also builds
  the basis for novel developments on the basis of the implemented reproducibility
  workflow as this paper is accompanied by a webpage that supports the impartial evaluation
  of the work by providing the data, code, and how this reports had been produced.\n"
title-meta: Busting the Myth! - Reproducibility in ANS Performance
---

@@

Air transportation is undergoing a fundamental transformation. To meet the future growth of air traffic, political decision makers, strategic planers, practitioners and researchers aim to enhance the performance of air navigation services (ANS).
At the time being there is no consistent and regionally harmonised deployment programme. Typically, the deployment programmes are revolving around public-private partnerships and regional funding programmes resulting in a wide mix of consortia and stakeholder subsets.
The actual deployment of novel concepts or technological enablers is then driven by the programmes and to a lesser extent by operational network wide priorities following a piece meal approach.

Moreover, the success of these deployments is typically supported by promising business cases and operational evaluations - often expressed as operational performance improvements. 
While these improvements are marketed, the underlying data and associated analyses in support of the results are seldomly made available. 
Explanations range from data sensitivity, commercial interests and property rights, through to complexity of the data analyses or simulations, and the associated volume of data.

Accordingly, ANS performance improvements are subject to a "credibility crisis" (varying levels of transparency) similar to reported problems in other domains, e.g. medicine [@kolata_2011], psychology [@ritchie_2012], ecology [@fraser_etal_2018]. Fidler and Wilcox (2018) provides a recent review of the reproducibility discussion [@fidler_wilcox_2018]. 
Part of the crisis sensation was driven by the diffusion of reported results as other researchers/analysts were able to discover substantial flaws in the data analyses. However, a crisis also offers room for self-correction.
Throughout the recent years, this lead to revisitng the concept of reproducible research.
This concept is based on the premises that an independent researcher/analyst is enabled to reproduce the reported results.
Reproducibility covers therefore a spectrum and mix of method, data, and analytical code/scripts per se openly available for scrutiny by others.

@@

Independent (operational) ANS performance builds on transparent and robust performance monitoring and reporting.
Within the European context, EUROCONTROL's Performance Review Unit (PRU) is committed to provide impartial analyses of the performance of the European Air Navigation System based on verified data.
With a view to enhance transparency, PRU is - inter alia - making performance-related data publicly available via its portal, [http://ansperformance.eu](http://ansperformance.eu), since early 2017 combined with the methodological documentation for the calculation of the performance indicators used under the EUROCONTROL Performance Revew System and the European Union Performance Scheme.
There has been recognition of using common and openly available tools and data in ATM research and operational validation.
With the advent of crowd sourced data and aviation databases there is an opportunity to advance the state of the art by combining open data and government organisation collected data to establish a curated platform supporting the quest for reproducibility in ANS Performance.
One of the associated building blocks is the performance reference trajectory [@spinielli_2018].
At the time of writing, the performance reference trajectory is in its design stage. The trajectory project builds on crowd sourced surveillance data fused with correlated position reports collected from European air navigation service providers.
Based on the Open Data model it will be possible to augment the data processing (e.g. data coverage, time synchronisation), establish derived data bases (e,g fleet data base), and derive associated performance models (e.g. aircraft performance model).
This setup enables the continual further development and validation of operational performance measures/indicators.

@@

This paper builds on the previous work and implements the reproducibility paradigm.
As a use-case example, a fully reproducible paper is devised assessing the performance in the arrival phase for 2 European airports. 
The underlying data, supporting data transformation and analysis, and the paper "source text" are made openly available at the paper companion web repository [https://github.com/euctrl-pru/reproducible-ans-performance-paper](https://github.com/euctrl-pru/reproducible-ans-performance-paper).
This enables the independent review and validation of the work reported.
Conceptually, an interested researcher/analyst can reproduce this paper by using the same toolset or validate the results by using the source data or interim data products if the complete reproduction of the study would be too time/resource consuming.

```{r}
#- Purpose and significance of the paper 
#- contribution of paper
```

The contributions of this paper are as follows:

* first fully reproducible ANS Performance related paper: The paper is written in RMarkdown allowing for the analysis of how text, code chunks, and visualisations are produced (including associated parameter settings) on the basis of the "source text".
* transparent data analyses and result generation through supporting web-page: All data used in this paper, intermediate data artifacts, and results including the associated transformation are publicly available for inspection, validation, or reproduction. 
* use-case analysis of ANS Performance during the arrival phase for 2 European airports: As an example the classical 'additional ASMA time' performance indicator is implemented on the basis of the provided data with showcasing trajectory derived capabilities to further study the efficiency of the arrival phase (e.g. impact of sequencing strategies [stack holdings, point merge]).

<!-- note: make sure that Section I does not significantly exceed first page -->

@@

Reproducibility has also to be seen in light of today's "publish or perish" context. It appears that higher emphasis is given to the frequency of publication than the validity of the produced results. Within the medical world cases like the Duke ovarian cancer scandal [@kolata_2011] are not single incidents, but also demonstrate how errors during the data analysis - in this case during the data preparation phase using spread sheets - can discredit the work.
Despite the increasing awareness, the adaption of reproducibility or moving to other open scholarship/investigation techniques is not yet achieved. 
This can be contrasted by emerging practices like data journalism. One interesting example is buzzFeed's article on the use of surveillance planes operated by law enforcement and military. The authors were able to demonstrate based on Open Data that these purposes range from tracking drug trafficking to testing new surveillance technology [@peteraldhous_2016]. The credibility of the article comes from the fact that all artifacts are made available to reproduce the findings or study the approach and data transformations leading to the conclusions of the article.

@@

* archaeology: Marwick et al. (2017) focusses on computational reproducibility in archaeological research by providing basic principles. The case study of their implementation uses open source software (R/Rmarkdown/git approach)  [@marwick_2017]. The paper advocates the publishing (analytical) data in support of the paper. This should not be an issue, even if there are constraints on the overall study data.
* engineering / signal processing: Vandewalle (2009) discuss the motivation and approach to reproducibility in signal processing (i.e. what, why, and how). Also here the case is made to open access to the code and data for validation purposes [@vandewalle_2009].
* operational ANS Performance monitoring: Spinielli et al. (2018) highlight the development and use of Open Flight Trajectories for reproducible ANS [@spinielli_2018].

## Practical Challanges

The reproducibility discussion focuses strongly on the recreation of results by other researchers/analysts.
A related problem is the re-generating of already completed (and potentially published) results with the - previously - available same input data. 
This is a fundamental issue that can be summarised as "can a future me re-do what I had done in the past"?
This problem space ranges across the complete data analysis life-cyle, including archiving the orignially used input data or the steps to generate the study data, documenting the data cleaning and transformation steps including the parameters chosen, etc.
By definition such a capability is a precondition for reproduction or replication by others. 
This implies that the analytical steps including the steps to produce summaries or estimates, visualisations, and associated conclusions are documented and archived. 
In order to enable others to recreate the own findings, such steps and documentation would need to be shareable.
This consideration has a strong bearing on the toolset chosen for this paper.

<!-- check if this can be merged with above -->
Some of the reasons why researchers/analysts do not publish data and computational procedures that allow reproduction will be hard to change [@daniel_nust_2018_2361304]. The predominant technical reasons have been listed above. However, there are also personal reasons such as the fear for floating ideas or loss of competitive advantage. Other reasons deal with more technical problems, e.g. lack of capabilities, procedures.

Another incentive for the adoption of an open and reproducible research practice can be seen in the wider recognition through increased citations, media attention, discussion with the community and collaborators, and ultimately funding opportunities.
The recent years have seen a strong use of research blogs and community platforms supporting the publication of smaller steps or intermediate results. The combination of web presentation and discussions offers a good basis for an initial dialogue.
Platforms like google scholar or researchgate.net offer additional features for outreach and communication.
This equally applies for paper archives such as arxiv.org for pre-prints.
Based on the credibility crisis discussion a number of journals are now promoting - or even requiring - the provision of analytical data in support of the paper publication. 
There is also an increasing number of requirements for national or European funded research with respect to data management and Open Data access to the research.

@@

Across the literature, researchers have identified a "reproducibility spectrum" ranging from the extreme, i.e. publication only (advertising and marketing of the done work), to almost or even fully replicable research.
The work of Marwick et al. [@marwick_2017] and Peng [@Peng1226] uses the analogy of a spectrum that can be mapped against the scoring of Vandewalle (2009) describing the level of reproducibility on a 0 to 5 scale [@marwick_2017].
In light of [@leek_2015], this paper conceptualises reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline" and accordingly "replicability of a study is the chance that an independent experiment targeting the same scientific question will produce a consistent result".

With Fig. \@ref(fig:reproducible-research) and the discussion in this section, this paper implements a fully reproducible workflow.
This is achieved by establishing this paper with open source software which allows to combine text, code, and visualisations.
The associated source text can also easily be read with a simple text editor/viewer. From that perspective, the text is fully inspectable.
This includes the majority of visualisations which are generated based on the data. Accordingly, the parameters and settings chosen can be validated by an independent researcher/analyst. Last but not least, the data for this paper is made available in conjunction with this text as Open Data. This also includes intermediate data artifacts which allows for the reproduction of the complete work or portions of it.
The conceptual approach and its building blocks will be presented in more detail in the next section.

@@

Analytical data refers to the data sets established based on the aforementioned processing code.
An essential defining characteristic of analytic data is that it is the data basis on which the reported results can be produced.
The abstraction between source data and analytic data allows for the sharing of such data as limitations are typically only appliable to the source data. We follow the notion of [@marwick_2017] by referring to the provision of a representative sample of 8 days accompanying this paper as source data.

@@

This work builds on the RStudio tools for the R language [@rcoreteam_2018] including Git (and the web-based repository managers GitHub and GitLab) as underlying version control system.
The R language  was originally developed within the statistical community supporting the task of statistical reporting by providing routines for the statistical computing and visualisation.
Being open source, the R community is actively engaging and sharing related software packages to enhance the core functionality.
Without limiting the impact of other packages, the development of knitr [@R-knitr] and RMarkdown [@xie2018], ggplot [@wickham_2016] for visualisation, and the so-called tidyverse packages and RStudio IDE [@rstudioteam_2015] represent an open source ecosystem for data analysis.

A key feature for the implementation of the reproducibility paradigm is the fact that RMarkdown supports the combination of text, coding, and visualisations in a single document. 
The latter can be rendered into various output formats, e.g. classical paper report-styles (pdf, MS Word), html web-pages, and slides (both classical Powerpoint style or html-based presentations).
The fact that outputs can be rendered to html pages also supports the embedding of interactive graphics. This also facilitates the communication and outreach amongst researchers/analysts as standard web technology becomes a standard outlet format.
It follows that changes to the code (and associated visualisations) or text can be readily updated by re-rendering the adapted report. 
This allows to develop standard templates for recurring reports based on varying input data, but it also serves the iterative development of larger reports (and studies).

A minimal example of an RMarkdown document, a plain-text file with the conventional
extension `.Rmd` is shown below:

@@

There are three components of an R Markdown document: the metadata, the text and the code.
The metadata, written between the pair of three dashes `---`, helps in the production of
output document (title, type of document [web page, PDF, MS Word], ...)
The text, which follows the metadata, is witten using Markdown, a simplified formatting syntax
which can include prose and code.
There are two types of (computer) code:

- A code chunk starts with three backticks and `r` indicates the
  language name,^[R Markdown supports 50 other languages.] and ends with
  three backticks.

@@

```{r rmd-example,echo=FALSE,fig.cap='(ref:rmd-example)',fig.pos='htb',fig.height=3}
knitr::include_graphics('figures/rmd-example.png', dpi = NA)
```

A particular strong point is that the R/RStudio ecosystem is open source software.
This includes also its extensibility with other packages. There exists a wide variety of packages for data analytical tasks and techniques. This mechanism can also be useful for the communication of results. Researchers/analysts are able to provide their code and supporting documentation as a package for other researchers to build on.
For example, throughout the last year, PRU has developed certain utility functions for reoccurring tasks of the data analytical pipeline.
It is now planned to wrap these functions into a package saving other researchers/analysts of recoding certain steps.
For this paper, the functions and data transformations are reported as part of the RMarkdown and supporting R files.
All these files are available via the project webpage.

@@

Throughout the recent year PRU developed the data analytical capabilities to fuse correlated position reports and ADSB surveillance data.
This work has been reported in [@koelle_open_2017] and [@spinielli_2018]. The idea is to establish a pan-European air situation picture openly available to research and practitioners.
Conceptually, the fused performance reference trajectory will be a key milestone in combatting the "marketing myths."
For example, operational improvements can be directly assessed by interested analysts as access to the Open Data enables the validation of the observable performance improvements. This allows for the application of commonly accepted methods and models to express performance benefits and  - ultimately - challenge related publications on deployments and achieved improvements.

At the end of 2018 the initial development project phase was closed and PRU is now investigating options for deploying and maintaining the processing modules on a cloud platform.
The goal is to establish and operate the underlying data processing capabilities throughout 2019 and launch the regular provision of the performance reference trajectory as Open Data in 2020.
It is also planned to establish a network and complement the air situation picture by deploying additional sensors or determining non-ADSB derived positions as a community effort.
As a mid-term goal associated data bases and models for performance shall be established and jointly maintained.

<!-- not word by word - but the info is contained elsewhere 
## Application Use Cases: ASMA, Point Merge, holdings

This paper revolves around advancing the state of the art by demonstrating the implementation of the reproducibility paradigm.
Accordingly, this paper builds on the concepts behind Fig. \@ref(fig:reproducible-research) and Fig. \@ref(fig:concept-approach).
This includes providing the data, code, and results.
All of this is version controlled and archived on <<LINK TO ARCHIVE OR WEB PAGE>>.

The approach and processing pipeline described in this paper builds on several data sets.
...

As a practical application use case, this paper replicates the additional ASMA time algorithm.
... describe AMSA here ... which data sets used ... how compared ... point to next section for details of this.

In a second step, the ASMA principle is moved to a distance based measure.
... describe what will be done ... how does this link to what we want to achieve ... point to next section for more details.

As examples for further developments based on the aforementioned capabilities, this paper presents a closer look at arrival flow management strategies.

-->

@@

@@

```{r data-preparation-pipeline,fig.cap='(ref:data-preparation-pipeline)',fig.pos='htb',fig.height=4}
knitr::include_graphics("figures/data-preparation-pipeline.png")
```

The raw data sources for the use cases of this paper have been processed by a pipeline of R scripts.
For example, the additional ASMA time indicator and the anticipated further breakdown of the arrival phase can be described by a set of tranistion points:

Figure \@ref(fig:data-preparation-pipeline) shows how the transition points 
(40 NM, entry/exit from holding/point-merge area, landing) for the use case airports
have been extracted.
Due to the size of the daily reference trajectory files, the pipeline operates
on one date at the time.
Assuming the relevant RT flight and position files are downloaded
from GCP as described in [@performancereviewunit_2019],
the steps and the relevant scripts (available in the companion repository for this paper) used are:

1. the script `R/extract-arrivals.R` extracts arrivals to the `APT` airport and generates:

    * a file with flight details
    * a file with all the arrivals' positions (`raw` trajectories)
    * a file with all the arrivals' positions (smoothed trajectories)
    * a file with all (smoothed) arrivals' positions within 40 NM of the airport's reference position (ARP)

@@

(ref:eidw-arrival-chart-east) Excerpt of Irish AIP showing point merge arc layout East of Dublin Airport.

```{r eidw-arrival-chart-east,fig.cap='(ref:eidw-arrival-chart-east)',fig.pos='hbt',fig.height=4}
knitr::include_graphics("figures/eidw-arrival-chart-east.png")
```

  This stage makes use of data extracted/extrapolated from the relevant AIP, see
  for example Fig. \@ref(fig:eidw-arrival-chart-east), such
  as ARP, runway name and threshold position/elevation, polygons for holding
  stacks or point-merge areas. Specifically the polygons have been heuristically defined from the
  point merge points in the AIP and enlarged by 2NM in order to cater for the spread of real flown
  trajectories,
  see the scripts `R/egll-data.R` and `R/eidw-data.R`.

@@

* aircraft/flight info (ID, callsign, registration, ADEP/ADES, ICAO 24 bit address)
* 4D position (timestamp $T$, longitude, latitude, altitude)
* flown distance $D$ from aerodrome of departure (ADEP)
* sequence number
* point info (sequence number, type of transition point [`P40`, `PHOLD`, `PLAND`])
* distance from ARP

@@

The previous sections described the data preparatory action.
For this paper, a reproducible data set for the analysis of arrival management techniques at London Heathrow and Dublin airport have been extracted. 
The data set comprises flights for the period of 1.-8. August 2017.
It can be inferred from the traffic mix at London Heathrow (major international hub) that only a limited subset of flights is not equipped with ADSB, while a slightly higher level of non-equipped flights is expected at Dublin (higher share of intra-European flights).

@@

ggplot(data = tmp
       ,mapping = aes(x = factor(DOF), fill = SOURCE)
       ) +
  geom_bar(position = "dodge") + 
  theme_bw() +
  facet_wrap(. ~ ADES) +
  theme(
    #legend.position = "bottom"
    legend.position = c(1, 0.65), legend.justification = c(1, 0)
    , axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)
    ) +
  labs(x = "date of flight"
       ,fill = "data source") 
```

Fig. \@ref(fig:coverage-egll) shows the overlap of the data between the arrivals extracted from the airport operator data flow (APDF) and the reference trajectory (RTRJ) for London Heathrow (EGLL) and Dublin (EIDW).

@@

For the time horizon of this paper, the coverage of the reference trajectory derived data is for Heathrow (EGLL) `r round(cov_perc$avg[cov_perc$ADES == "EGLL"], 2) * 100`% and for Dublin `r round(cov_perc$avg[cov_perc$ADES == "EIDW"], 2) * 100`%.

@@

The reference time $i$ is determined for flights with similar arrival characteristics (c.f. [http://ansperformance.eu/references/definition/additional_asma_time.html](http://ansperformance.eu/references/definition/additional_asma_time.html)). These include 1.) the arrival entry sector at 40NM from the airport, 2.) the aircraft weigth turbulence category and engine type, and 3.) the actual landing runway.
The reference times are built for a longer time horizon and extracted from the PRU data base as additional source data.

@@

As an example, the distribution of the actual arrival transit times for arrivals into Heathrow (EGLL) are shown in Fig. \@ref(fig:ASMA-travel-egll). Heathrow operates holding stacks to tactically optimise the arrival throughput.
Dependent on the traffic situation, aircraft may experience longer holding times (i.e. long tails of boxplots in Fig. \@ref(fig:ASMA-travel-egll)).

@@

Fig. \@ref(fig:ASMA-compare) compares the additional ASMA time extracted from the PRU performance dashboard, the trajectory based calculation with the dashboard derived reference time, and a 20th percentile approach purely based on the trajectory data.

@@

knitr::kable(dist_asma, digits = 2
             ,caption = "Average additional distance flown (15th pct)")
```

Tab. \@ref(tab:tbl-asma-dist) lists the determined average additional distance at the study airports.
The total additional distance is determined as the aggregated sum of the differences in each segmnent
with 15-th percentile reference distance.

@@

source(here::here("R/plot-eidw-tma.R"))

>>>>

---
title: Towards Reproducibility in ANS Performance
author:
- Enrico Spinielli
- Rainer Koelle
date: "`r format(Sys.time(), '%a %b %d %Y @  %X')`"
output:
  bookdown::pdf_book:
    base_format: rticles::ieee_article
    fig_caption: no
    includes:
      in_header: header.tex
      keep_tex: yes
    number_sections: yes
    toc: no
    with_dblfloatfix: no
    with_subfig: no
  html_document: default
bibliography: references.bib
color-links: yes
csl: ieee-transactions-on-aerospace-and-electronic-systems.csl
affiliation:
  institution-columnar: true
  institution:
  - author:
    - email: enrico.spinielli@eurocontrol.int
      name: Enrico Spinielli
    - email: rainer.koelle@eurocontrol.int
      name: Rainer Koelle
    department: Performance Review Unit
    location: Brussels, Belgium
    name: EUROCONTROL
keywords:
- reproducibility
- ANS performance
- flight trajectory
- open data
- performance indicator
link-citations: yes
author-meta: Enrico Spinielli, Rainer Koelle
papersize: a4paper
abstract: "Air transportation is undergoing a fundamental transformation. On the political,
  strategic, and operational level work is directed to meet the future growth of air
  traffic.\nTo meet this goal, operational concepts and technological enablers are
  deployed with the promises of distinct performance improvements. These improvements
  are typically widely marketed, however, the underlying evidence is not made available.
  From that respect Air Navigation System Performance faces varying levels of transparency
  as seen in other industries. \nThis paper implements the reproducibility paradigm
  by providing data, code, and results openly to enable another researcher or analyst
  to reproduce - and potentially - improve or build on our work. This work is part
  of a wider thread to establish Open Data and Open Source based data analytical capabilities
  for Air Navigation System Performance in Europe.\nThe implementation of the paradigm
  is demonstrated on the basis of the performance reference trajectory by replicating
  a classical performance measure, and then moving into investigating the application
  of trajectory based information for enhancing the current state of performance monitoring
  for the arrival phase. This paper applies the concept and approach as a use case
  analysis for 2 European airports with different strategies for sequencing arrival
  traffic (i.e. stack holding and point merge).\nThe research reported demonstrates the
  feasibility of the reproducibility approach. This work introduces a novel workflow by making available a
  companion web repository that provides data, code, and instructions to fully reproduce the
  paper.\n"
title-meta: Towards Reproducibility in ANS Performance
---

@@

To meet the future growth of air traffic, political decision makers, strategic planers, practitioners and researchers aim to enhance the performance of air navigation services (ANS).
Typically, deployment programmes are revolving around public-private partnerships and regional funding programmes resulting in a wide mix of consortia and stakeholder subsets.
The actual deployment of novel concepts or technological enablers is then driven by the programmes and to a lesser extent by operational network wide priorities following a piece meal approach.

Moreover, the success of these deployments is typically supported by promising business cases and operational performance improvements. 
While these improvements are marketed, the underlying data and associated analyses in support of the results are seldomly made available. 
Explanations range from data sensitivity, commercial interests and property rights, through to complexity of the data analyses or simulations, and the associated volume of data.
Similar practise is reported in other domains, e.g. medicine [@kolata_2011], psychology [@ritchie_2012], ecology [@fraser_etal_2018]. Fidler and Wilcox [@fidler_wilcox_2018] provides a recent review of the reproducibility discussion. 
Throughout the recent years, this lead to revisiting the concept of reproducible research.
This concept is based on the premises that an independent researcher/analyst is enabled to reproduce the reported results.
Reproducibility covers therefore a spectrum and mix of method, data, and analytical code/scripts per se openly available for scrutiny by others.

@@

Independent (operational) ANS performance builds on transparent and robust performance monitoring and reporting.
Within the European context, EUROCONTROL's Performance Review Unit (PRU) is committed to provide impartial analyses of the ANS performance.
With a view to enhance transparency, PRU is - inter alia - making performance-related data publicly available via its portal, [http://ansperformance.eu](http://ansperformance.eu), since early 2017 combined with the methodological documentation for the calculation of the performance indicators used under the EUROCONTROL Performance Review System and the European Union Performance Scheme.
There has been recognition of using common and openly available tools and data in ATM research and operational validation.
With the advent of crowd sourced data and aviation databases there is an opportunity to advance the state of the art by combining open data and government organisation collected data to establish a curated platform supporting the quest for reproducibility in ANS Performance.
One of the associated building blocks is the performance reference trajectory [@spinielli_2018].
At the time of writing, the performance reference trajectory is in its design stage. The trajectory project builds on crowd sourced surveillance data fused with correlated position reports collected from European air navigation service providers.
Based on the Open Data model it will be possible to augment the data processing (e.g. data coverage, time synchronisation), establish derived data bases (e,g fleet data base), and derive associated performance models (e.g. aircraft performance model).
This setup enables the continual further development and validation of operational performance measures/indicators.

@@

This paper builds on the previous work and implements the reproducibility paradigm.
As a use-case example, a fully reproducible paper is devised assessing the performance in the arrival phase for 2 European airports. 
The underlying data, supporting data transformation and analysis, and the paper "source text" are made openly available at the paper companion web repository [https://github.com/euctrl-pru/reproducible-ans-performance-paper](https://github.com/euctrl-pru/reproducible-ans-performance-paper).
This enables the independent review and validation of the work reported.

@@

Reproducibility has also to be seen in light of today's "publish or perish" context. It appears that higher emphasis is given to the frequency of publication than the validity of the produced results. Within the medical world cases like the Duke ovarian cancer scandal [@kolata_2011] are not single incidents, but also demonstrate how errors during the data analysis - in this case during the data preparation phase using spread sheets - can discredit the work.
Despite the increasing awareness, the adoption of reproducibility or moving to other open scholarship/investigation techniques is not yet achieved. 
This can be contrasted by emerging practices like data journalism. One interesting example is buzzFeed's article on the use of surveillance planes operated by law enforcement and military. The authors were able to demonstrate based on Open Data that these purposes range from tracking drug trafficking to testing new surveillance technology [@peteraldhous_2016]. The credibility of the article comes from the fact that all artefacts are made available to reproduce the findings or study the approach and data transformations leading to the conclusions of the article.

@@

* archaeology: Marwick et al. (2017) focusses on computational reproducibility in archaeological research by providing basic principles. The case study of their implementation uses open source software (R/RMarkdown/git approach)  [@marwick_2017]. The paper advocates the publishing (analytical) data in support of the paper. This should not be an issue, even if there are constraints on the overall study data.
* engineering / signal processing: Vandewalle [@vandewalle_2009] discuss the motivation and approach to reproducibility in signal processing (i.e. what, why, and how). Also here the case is made to open access to the code and data for validation purposes .
* operational ANS Performance monitoring: Spinielli et al. [@spinielli_2018] highlight the development and use of Open Flight Trajectories for reproducible ANS .

@@

Across the literature, researchers have identified a "reproducibility spectrum" ranging from the extreme, i.e. publication only (advertising and marketing of the done work), to almost or even fully replicable research.
The work of Marwick et al. [@marwick_2017] and Peng [@Peng1226] uses the analogy of a spectrum that can be mapped against the scoring of Vandewalle [@vandewalle_2009] describing the level of reproducibility on a 0 to 5 scale [@marwick_2017].
In light of [@leek_2015], this paper conceptualises reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline" and accordingly "replicability of a study is the chance that an independent experiment targeting the same scientific question will produce a consistent result".

With Fig. \@ref(fig:reproducible-research) and the discussion in this section, this paper implements a fully reproducible workflow.
This is achieved by establishing this paper with open source software which allows to combine text, code, and visualisations.
The associated source text can also easily be read with a simple text editor/viewer. From that perspective, the text is fully inspectable.
This includes the majority of visualisations which are generated based on the data. Accordingly, the parameters and settings chosen can be validated by an independent researcher/analyst. Last but not least, the data for this paper is made available in conjunction with this text as Open Data. This also includes intermediate data artefacts which allows for the reproduction of the complete work or portions of it.
The conceptual approach and its building blocks will be presented in more detail in the next section.

@@

Analytical data refers to the data sets established based on the aforementioned processing code.
An essential defining characteristic of analytic data is that it is the data basis on which the reported results can be produced.
The abstraction between source data and analytic data allows for the sharing of such data as limitations are typically only applicable to the source data. We follow the notion of [@marwick_2017] by referring to the provision of a representative sample of 8 days accompanying this paper as source data.

@@

This work builds on the RStudio tools for the R language [@rcoreteam_2018] including Git (and the web-based repository managers GitHub and GitLab) as underlying version control system.
The R language  was originally developed within the statistical community supporting the task of statistical reporting by providing routines for the statistical computing and visualisation.
Being open source, the R community is actively engaging and sharing related software packages to enhance the core functionality.
Without limiting the impact of other packages, the development of knitr [@R-knitr] and RMarkdown [@xie2018], ggplot [@wickham_2016] for visualisation, and the so-called tidyverse packages and RStudio IDE [@rstudioteam_2015] represent an open source ecosystem for data analysis.
A key feature for the implementation of the reproducibility paradigm is the fact that RMarkdown supports the combination of text, analytical code, and visualisations in a single document. 
A minimal example of an RMarkdown document, a plain-text file with the conventional
extension `.Rmd` is shown below:

@@

There are three components of an R Markdown document: the metadata, the text and the code.
The metadata, written between the pair of three dashes `---`, helps in the production of
output document (title, type of document [web page, PDF, MS Word], ...)
The text, which follows the metadata, is written using Markdown, a simplified formatting syntax
which can include prose and code.
There are two types of (computer) code:

- A code chunk starts with three backticks and `r` indicates the
  language name,^[R Markdown supports 50 other computer languages.] and ends with
  three backticks.

@@

```{r rmd-example,echo=FALSE,fig.cap='(ref:rmd-example)',fig.pos='htb',fig.height=2}
knitr::include_graphics('figures/rmd-example.png', dpi = NA)
```

@@

Throughout the recent year PRU developed the data analytical capabilities to fuse correlated position reports and ADSB surveillance data.
This work has been reported in [@koelle_open_2017] and [@spinielli_2018]. The idea is to establish a pan-European air situation picture openly available to researchers and practitioners.
Conceptually, the performance reference trajectory will be a key milestone in combatting the "marketing myths."
For example, operational improvements can be directly assessed by interested analysts as access to the Open Data enables the validation of the observable performance improvements. This allows for the application of commonly accepted methods and models to express performance benefits and  ultimately challenge related publications on deployments and achieved improvements.

At the end of 2018 the initial development project phase was closed and PRU is now investigating options for deploying and maintaining the processing modules on a cloud platform.
The goal is to establish and operate the underlying data processing capabilities throughout 2019 and launch the regular provision of the performance reference trajectory as Open Data in 2020.

@@

The raw data sources for the use cases of this paper have been processed by a pipeline of R scripts.
For example, the additional ASMA time indicator and the anticipated further breakdown of the arrival phase can be described by a set of transition points:

@@

```{r data-preparation-pipeline,fig.cap='(ref:data-preparation-pipeline)',fig.pos='htb',fig.height=2.5}
knitr::include_graphics("figures/data-preparation-pipeline.png")
```

Figure \@ref(fig:data-preparation-pipeline) shows how the transition points 
(40 NM, entry/exit from holding/point-merge area, landing) for the use case airports
have been extracted.
The following steps and scripts comprise [@performancereviewunit_2019]:

1. the script `R/extract-arrivals.R` extracts arrivals to the `APT` airport and generates files with:

    * flight details
    * arrivals' positions (`raw`  and smoothed trajectories)
    * (smoothed) arrivals' positions within 40 NM of the airport's reference position (ARP)

@@

  This stage makes use of data extracted/extrapolated from the relevant AIP such
  as ARP, runway name and threshold position/elevation, polygons for holding
  stacks or point-merge areas. Specifically the polygons have been heuristically defined from the
  point merge points in the AIP and enlarged by 2NM in order to cater for the spread of real flown
  trajectories,
  see the scripts `R/egll-data.R` and `R/eidw-data.R`.

@@

* aircraft/flight info: ID, callsign, registration, ADEP/ADES, ICAO 24 bit address;
* 4D position: timestamp $T$, longitude, latitude, altitude;
* position info: flown distance $D$ from aerodrome of departure (ADEP), sequence number,
  distance to ARP as a proxy of aerodrome of destination (ADES);
* point info: sequence number, type of transition point [`P40`, `PHOLD`, `PLAND`]

@@

For this paper, a reproducible data set for the analysis of arrival management techniques at London Heathrow and Dublin airport have been extracted. 
The data set comprises flights for the period of 1.-8. August 2017.
It can be inferred from the traffic mix at London Heathrow (major international hub) that only a limited subset of flights is not equipped with ADSB, while a slightly higher level of non-equipped flights is expected at Dublin (higher share of intra-European flights).

@@

ggplot(data = tmp
       ,mapping = aes(x = factor(DOF), fill = SOURCE)
       ) +
  geom_bar(position = "dodge") + 
  theme_bw() +
  facet_wrap(. ~ ADES) +
  theme(
    #legend.position = "bottom"
    legend.position = c(0.95, 0.60), legend.justification = c(1, 0)
    , axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)
    ) +
  labs(x = "date of flight"
       ,fill = "data source") 
```

@@

Fig. \@ref(fig:coverage-egll) shows the overlap of the data between the arrivals extracted from the airport operator data flow (APDF) and the reference trajectory (RTRJ) for London Heathrow (EGLL) and Dublin (EIDW).
For the time horizon of this paper, the coverage of the reference trajectory derived data is for Heathrow (EGLL) `r round(cov_perc$avg[cov_perc$ADES == "EGLL"], 2) * 100`% and for Dublin `r round(cov_perc$avg[cov_perc$ADES == "EIDW"], 2) * 100`%.

@@

The reference time $i$ is determined for flights with similar arrival characteristics (c.f. [http://ansperformance.eu/references/definition/additional_asma_time.html](http://ansperformance.eu/references/definition/additional_asma_time.html)). These include 1.) the arrival entry sector at 40NM from the airport, 2.) the aircraft weight turbulence category and engine type, and 3.) the actual landing runway.
The reference times are built for a longer time horizon and extracted from the PRU data base as additional source data.

@@

As an example, the distribution of the actual arrival transit times for arrivals into Heathrow (EGLL) are shown in Fig. \@ref(fig:ASMA-travel-egll). Heathrow operates holding stacks to tactically optimise the arrival throughput.
Dependent on the traffic situation, aircraft may experience longer holding times (i.e. long tails of box plots in Fig. \@ref(fig:ASMA-travel-egll)).

@@

Fig. \@ref(fig:ASMA-compare) compares the additional ASMA time extracted from the PRU performance dashboard, the trajectory based calculation with the dashboard derived reference time, and a 20-th percentile approach purely based on the trajectory data.

@@

knitr::kable(dist_asma, digits = 2
             ,caption = "Average additional distance flown (15-th pct)")
```

@@

library(dplyr)
library(sf)
library(ggplot2)
library(units)
